% Copyright (c) 2012 Taylor Hutt, Logic Magicians Software
%
% This program is free software: you can redistribute it and/or
% modify it under the terms of the GNU General Public License as
% published by the Free Software Foundation, either version 3 of the
% License, or (at your option) any later version.
%
% This program is distributed in the hope that it will be useful, but
% WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
% General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program.  If not, see <http://www.gnu.org/licenses/>.
%
\chapter{Errors}

\section{E1000}

\begin{verbatim}
E1000: Invalid configuration: 'load-configuration-function' not set
\end{verbatim}

You have loaded \lmsbw with a configuration file, but you have not set
the \texttt{load-configuration-function} configuration value.  See
\xref{aa:load-configuration-function} for details on setting this value.

\section{E1001}

\begin{verbatim}
E1001: Invalid configuration: 'load-configuration-function' \
       must be defined in an included Makefile
\end{verbatim}

The function referenced by your \texttt{load-configuration-function}
value does not exist.  See \xref{aa:load-configuration-function} for
details on properly setting this value.

\section{E1002}

\begin{verbatim}
E1002: Configuration file '<pathname>' redefines component
 '<component>' that is already declared in '<pathname>'
\end{verbatim}

Each component name must be globally unique to a project.  You have
two configuration files declaring the same component name.

\section{E1003}

\begin{verbatim}
E1003: Reason '<reason>' is not 'build' nor 'image'
\end{verbatim}

You have attempted to use an invalid \texttt{reason}.  See
\xref{variables:reason} for details on the allowed values for this
variable.

\section{E1004}

\begin{verbatim}
E1004: Component '<component>' cannot have itself as a prerequisite
\end{verbatim}

\begin{quote}
  \textbf{I am my own grandfather}

  I married a widow who has a grown-up daughter.  My father, who
  visited us quite often, fell in love with my step-daughter and
  married her.  Hence, my father became my son-in-law and my
  step-daughter became my mother.  Some months later, my wife gave
  birth to a son, who became the brother-in-law of my father, as well
  as my uncle. The wife of my father, my step-daughter, also had a
  son.

  \underline{Algorithm's + Data Structures = Programs}

  \emph{Niklaus Wirth}
\end{quote}

\section{E1005}

\begin{verbatim}
E1005: No verb '<verb>' for component '<component>'
E1005: Does this component exist?  Use 'lmsbw report'
\end{verbatim}

You have attempted to use a valid \lmsbw verb on a component that does
not exist.  You have probably misspelled the component name.

If you had used an invalid verb, \lmsbw will report that there is no
rule to make the target.

\section{E1007}

\begin{verbatim}
E1007: Component prerequisite path '<component-0> -> \
       (<component> ->)... <component-0>' produces a cycle
\end{verbatim}

The set of components and their prerequisites in \lmsbw must be a
directed acyclic graph (DAG).  The set of components shown in the
error message produce a graph with a cycle.

You must restructure your componnents so that they form a DAG.

\section{E1008}

\begin{verbatim}
E1008: Unknown component '<component>'
\end{verbatim}

You have attempted to use an unknown component.  This will typically
occur if you try to set an attribute or call a component declaration
function on a component which has not yet been declared.

To remedy this situation, make sure you always declare components
before you try to change them.

\section{E1009}

\begin{verbatim}
E1009: Function '<function>' for component '<component>' not written
\end{verbatim}

This is an internal error which should only be seen when an extension
to \lmsbw is incomplete.  Internally, \lmsbw will call functions which
are written to specifically process a component of a certain
\emph{kind} (\xref{variables:kind}).  If you have added a new
component \emph{kind} to \lmsbw, you have missed the implementation of
a necessary function.  You may use the \emph{source} verion of the
missing function as a reference on the function semantics.

