% Copyright (c) 2012 Taylor Hutt, Logic Magicians Software
%
% This program is free software: you can redistribute it and/or
% modify it under the terms of the GNU General Public License as
% published by the Free Software Foundation, either version 3 of the
% License, or (at your option) any later version.
%
% This program is distributed in the hope that it will be useful, but
% WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
% General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program.  If not, see <http://www.gnu.org/licenses/>.
%
\chapter{Overriding Default Rules} \label{chap:overriding}

Occasionally there will be a component \makefile that cannot be edited
for various reasons that does not conform to the \lmsbw requirements.

When \lmsbw's default for producing a component is insufficient or
when it's not feasible to modify the component's build process, it is
possible to override the built-in method for building \& installing by
providing particularly named targets in the component's configuration
file.

This chapter describes the process to override these aspects of
producing a component.

\section{Overriding Build}\label{overriding:overriding-build}

In normal practice, it should be exceedingly rare for a component
build process to be overridden; even though it should be rare, \lmsbw
still provides a method of doing just that.  Here are a few of the
reasons to override the build phase of a component:

\begin{itemize}
\item Test \bni Infrastructure by not building a component

  How does the \bni system function if a required component is not
  actually built?

  Will it fail during \emph{install}, or while integrating into a
  production deliverable?  Will it fail with an easy-to-diagnose
  error?

  The simple test would be to add a build override target that simply
  does nothing.

\item Test \bni Infrastructure by corrupting build

  Will the \bni system recognize if a build has been corrupted?
  Perhaps only some files are install, or maybe the executables are
  corrupt, or the executables are for a different architecture?

  To test this, override the build target, build the component (using
  the \lmsbw infrastructure) and then corrupt the build output
  in-situ.

\item Apply a patch in-situ

  The component's source tree cannot be changed in the source tree,
  but it is possible to apply a patch to the sources after they have
  been copied to the build directory.

  To do this, override the build rule to first apply the
  patches\footnote{When the overriden build rule is dependent upon
    \texttt{sync}, the source files will have been copied to the build
    directory before the rule is invoked.}, then use the \lmsbw
  infrastructure to invoke the component's build process.

\item Component may not need a build; only installs static files

  Perhaps the component doesn't have a build; maybe only static files
  are installed.  In this case, override the build to do nothing, and
  save a minute amount of time.

\item Other unknown reasons
\end{itemize}

Once it has been determined that overriding the automatic rules to
perform a component build is necessary, the actual work is quite
trivial; add a specially named \makefile target to the component
configuration file, and supply the new commands associated with the
target.  The name of the special target is based on the component
name; figure~\tabref{overriding:override-build-alpha} shows how this
is done for a component named \texttt{alpha}.

\begin{figure}[h]
  \hrulefill
\begin{verbatim}
component.build.alpha:
        echo "Overriding trampoline; not building alpha";
\end{verbatim}
  \caption{Overriding Build for Component \texttt{alpha}}\label{overriding:override-build-alpha}
  \hrulefill
\end{figure}


If the component configuration file does \emph{not} have this specific
target, then the default build rules of \lmsbw will take effect, and
tbe component's \makefile will be invoked to build the component.

To ultimately execute the component's actual build process, use the
same code of \lmsbw proper to initiate this, as shown in
figure~\tabref{overriding:invoke-build-like-lmsbw}.

\begin{figure}[h]
  \hrulefill
\begin{footnotesize}
\begin{verbatim}
$(MAKE)                                                          \
    $(DEFAULT_$(call uc,$(LMSBW_C_KIND))_COMPONENT_MAKE_OPTIONS) \
    $(LMSBW_C_BUILD_TARGET)
\end{verbatim}
\end{footnotesize}
  \caption{Invoking Build Like \lmsbw}\label{overriding:invoke-build-like-lmsbw}
  \hrulefill
\end{figure}


Be aware when overriding default targets that the override does not
provide any prerequisites.  When overriding the build target, for
example, without the proper prerequisites, it will not be possible to
use the code shown in
figure~\tabref{overriding:invoke-build-like-lmsbw} to build the
component.  See the \texttt{override} sample (\xref{samples:override})
for details, and look at \texttt{component.makefile} to determine the
necessary prerequisites.

\section{Overriding Install}\label{overriding:overriding-install}

Overriding the default install rules is a task that will happen more
frequently than overriding the build rules.  Here are some reaons why
overriding the default install rules might be necessary:

\begin{itemize}
\item Test \bni Infrastructure by not installing component
\item Component \makefile has no install rule, and cannot be modified
\item Component \makefile does not use \destdir
\item Component does not install all needed files
\item Other unknown reasons
\end{itemize}

Once it has been determined that overriding the automatic rules to
perform a component install is necessary, the actual work is quite
trivial; add a specially named \makefile target to the component
configuration file, and supply the new commands to associated with the
target.  The name of the special target is based on the component
name; figure~\tabref{overriding:override-install-alpha} shows how this
is done for a component named \texttt{alpha}.


\begin{figure}[h]
  \hrulefill
\begin{small}
\begin{verbatim}
component.install.alpha:
        $(MKDIR) --parents $(DESTDIR)/usr/include/alpha
        $(CP) --recursive                                  \
             $(LMSBW_C_BUILD_WORKING_DIRECTORY)/include/*  \
             $(DESTDIR)/usr/include/alpha
\end{verbatim}
\end{small}
  \caption{Overrding \texttt{install} For Component \texttt{alpha}}\label{overriding:override-install-alpha}
  \hrulefill
\end{figure}

If the component configuration file does \emph{not} have this specific
target, then the default install rules of \lmsbw will take effect, and
the component's \makefile will be invoked to install the component.

To ultimately execute the component's actual build process, use the
same code of \lmsbw proper to initiate this, as shown in
figure~\tabref{overriding:invoke-install-like-lmsbw}.

\begin{figure}[h]
  \hrulefill
\begin{footnotesize}
\begin{verbatim}
$(MAKE)                                                          \
    $(DEFAULT_$(call uc,$(LMSBW_C_KIND))_COMPONENT_MAKE_OPTIONS) \
    $(LMSBW_C_INSTALL_TARGET)
\end{verbatim}
\end{footnotesize}
  \caption{Invoking Component Install Like \lmsbw}\label{overriding:invoke-install-like-lmsbw}
  \hrulefill
\end{figure}


Be aware when overriding default targets that the override does not
provide any prerequisites.  When overriding the install target, for
example, without the proper prerequisites, it will not be possible to
use the code shown in
figure~\tabref{overriding:invoke-install-like-lmsbw} to install the
component.  See the \texttt{override} sample (\xref{samples:override})
for details, and look at \texttt{component.makefile} to determine the
necessary prerequisites.

