% Copyright (c) 2012 Taylor Hutt, Logic Magicians Software
%
% This program is free software: you can redistribute it and/or
% modify it under the terms of the GNU General Public License as
% published by the Free Software Foundation, either version 3 of the
% License, or (at your option) any later version.
%
% This program is distributed in the hope that it will be useful, but
% WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
% General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program.  If not, see <http://www.gnu.org/licenses/>.
%
\chapter{Toolchain Usage}\label{chap:toolchain-usage}

\lmsbw trivially facilitates using custom-built toolchains for
building a project: just specify the toolchain root directory
(\xref{usinglmsbw:toolchains-root}), and the name of a sub-directory
inside the root directory (\xref{usinglmsbw:toolchain}).

There are many reasons to use a custom-built toolchain over the one
present on the host computer.  The foremost reason for this is that it
becomes possible to reproduce any build of a project, at any time and
on any machine; creating a bit-for-bit identical build at some point
in the future using the toolchain on the host -- which may have been
upgraded -- is unlikely to be be successful.  Secondarily, everyone on
the development team will be building the exact same images,
regardless the default toolchain installed on their system.

\lmsbw has been designed to work with toolchains produced by the
crosstool-ng program (\xref{chap:toolchain-configuration}).  This has
several implications:

\begin{itemize}
\item Reproducible Builds

  In many industries -- industrial control, for example -- it's
  important to be able rebuild a project in exactly the same manner as
  it was built before.  If simply using the host's toolchain, upgrades
  of the compiler may no longer generate the same code, and the
  project may no longer work correctly.

  However, if using \lmsbw, because the toolchain has been
  specifically configured and built for the project, it can always be
  identically built with that toolchain.

\item Architecture Cross Compilation

  Allowing the ability to switch toolchains enables the ability to
  transparently cross-compile the program to a completely different
  architecture.

  Couple cross-compilation with the component configuration model of
  \lmsbw, and it is easy to have a project with a core engine that is
  present in every SKU, and an architecture-specific portion that is
  automatically included when necessary
  (\xref{wrap:master-configuration}).

\item{OS Version Cross Compilation}

  To compile for different versions of Linux, a different version of
  \texttt{glibc} will sometimes be needed .  Since the C library is
  part of the toolchain, this becomes very easy.

\item Different Toolchains for Components

  Sometimes a compiling a component with a different compiler is
  required.
\end{itemize}

\section{Sample Configurations}

\lmsbw provides several \ctng configurations to facilitate quickly
being able to cross compile projects.  In order to be able to use
these sample configurations, first configure \& install \ctng.  Please
follow the instructions in \xref{toolchain-config:prerequisites} and
\xref{toolchain-config:install-ctng}.

Once \ctng has been installed, it's very simple to install these
sample toolchains.

\begin{itemize}
  \item Change directory

    Select a toolchain to build, and change to that directory.  For
    example, to build one of the MIPS toolchains:

\begin{footnotesize}
\begin{verbatim}
cd ${LMSBW}/toolchains/mips/be_linux.3.3.4_gcc.4.6.3_eglibc.2.15
\end{verbatim}
\end{footnotesize}

  \item Build

    To build and install the toolchain, execute:

    \texttt{/opt/crosstool-ng/bin/ct-ng build}

    The toolchains will be placed in \texttt{/opt/toolchains}.
\end{itemize}


\section{Using a Custom Toolchain}

The following steps must be taken to use a toolchain:

\begin{enumerate}
\item Configure \& Build toolchains

  Follow the instruction in \xref{chap:toolchain-configuration} to
  build toolchains.

\item Specify toolchain root directory

  When executing \lmsbw, include the \texttt{--toolchains-root}
  argument.  See \xref{usinglmsbw:toolchains-root}.

\item Specify desired toolchain

  There are three levels at which the toolchain can be set for any
  project.

  \begin{itemize}
  \item component

    To indicate that a component must be built with a specific
    toolchain using the \texttt{component\_attribute\_toolchain}
    (\xref{api:toolchain}) API.

    Once a toolchain has been specified at the component level, the
    component will always be built using that toolchain.

    If the specified toolchain does not exist in the toolchain root
    directory, \lmsbw will produce an error.

    If a component does not have a toolchain specified, the global
    toolchain setting is used.

  \item global

    To use a specific toolchain to build all components with their
    \emph{reason} set to \texttt{image} (\xref{variables:reason}),
    when executing \lmsbw, include the \texttt{--toolchain} argument.

    See \xref{usinglmsbw:toolchain} for more details on this global
    argument.

    If the specified toolchain does not exist in the toolchain root
    directory, \lmsbw will produce an error.

    If a global toolchain is not specified, then the toolchain
    installed on the host computer is used.

  \item host

    If no toolchain has been specified, then \lmsbw will use the
    toolchain installed on the host computer.

    If no build tools are installed on the host, then components will
    fail to build; the necessary build tools must already be installed
    on the host computer.

  \end{itemize}

\end{enumerate}

\section{Examples of Using Toolchains}

An example of using various toolchains to build one of the sample
projects will helpful in understanding how easy it is to cross compile
to different architectures, or to just use a specific compiler for the
same architecture as the host.

We will be using the \texttt{lite-consumer-pro} sample for this
example.

\subsection{Reducing Repetition with Shell Functions}

To reduce the repetitive, keyboard-based, work for building the sample
with different toolchains, several \texttt{bash} functions will be
written and reused.

\begin{itemize}
  \item{\texttt{set\_tc}}

    This function is used to \emph{set the toolchain} that will be
    used to build the sample.  It also configures the product within
    the \texttt{lite-consumer-pro} sample that should be built.

\begin{verbatim}
function set_tc ()
{
  export PRODUCT=lite;
  export TOOLCHAIN=${1};
}
\end{verbatim}

\item{\texttt{build}}

  This function invokes \lmsbw to build the sample.  This can only be
  used after \texttt{set\_tc} has been called.

\begin{verbatim}
function build ()
{
  lmsbw
    --toolchains-root /opt/toolchains
    --toolchain ${TOOLCHAIN}
    --build-root /tmp/lite
    --configuration product.cfg $*;
}
\end{verbatim}

\item{\texttt{check\_lite\_file}}

  This function executes \texttt{file} on the file produced by the
  sample's build process.  It obtains the install directory of the
  file by using the \texttt{installdir} (\xref{usinglmsbw:installdir})
  and discarding the first field -- the component name.

\begin{small}
\begin{verbatim}
function check_lite_file ()
{
  file $(build installdir|cut -f 2 -d ':')/usr/bin/lite
}
\end{verbatim}
\end{small}

\item{\texttt{pt}}

  This function performs the actual test by wrapping all the previous
  functions together into a single command.  This command will be used
  to build the sample for several toolchains.

\begin{verbatim}
function pt ()
{
  local tc=${1};
  shift;
  set_tc ${tc};
  build
  check_lite_file
}
\end{verbatim}
\end{itemize}

\subsection{Little Endian Arm}
The commands and results for a \emph{little endian Arm} toolchain can be
seen in figure \tabref{toolchain-usage:le-arm}.

\subsection{x86-64}

The commands and results for an \emph{x86-64} toolchain can be seen in
figure \tabref{toolchain-usage:x86-64}.

\subsection{Big Endian MIPS}
The commands and results for a \emph{big endian MIPS} toolchain can be
seen in figure \tabref{toolchain-usage:be-mips}.

\subsection{Little Endian MIPS}

The commands and results for a \emph{little endian MIPS} toolchain can
seen in figure \tabref{toolchain-usage:le-mips}.

\begin{landscape}
\begin{figure}
\hrulefill
\begin{verbatim}
pt arm-le_linux.3.3.4_gcc.4.6.3_eglibc.2.15-linux-gnueabi

lite: Creating directory: '/tmp/lite/target/lite/47145117a11a5ce831079351eec2c54d/build'
lite: Creating directory: '/tmp/lite/target/lite/47145117a11a5ce831079351eec2c54d/destdir'
Creating directory: '/tmp/lite/target/install/451336eb87a4169e325952ca674c69db'
lite: Trampoline to 'lite' build system
lite: Updating manifest
lite: Install
/tmp/lite/target/install/451336eb87a4169e325952ca674c69db/usr/bin/lite: ELF 32-bit LSB
   executable, ARM, version 1 (SYSV), dynamically linked (uses shared libs),
   for GNU/Linux 3.3.4, not stripped
\end{verbatim}
\caption{Little Endian Arm}\label{toolchain-usage:le-arm}
\hrulefill
\end{figure}

\begin{figure}
\hrulefill
\begin{verbatim}
pt x86_64-64_linux.3.3.4_gcc.4.6.3_eglibc.2.15-linux-gnu

lite: Creating directory: '/tmp/lite/target/lite/160a8847f104d5b6f5432b98dfd6e319/build'
lite: Creating directory: '/tmp/lite/target/lite/160a8847f104d5b6f5432b98dfd6e319/destdir'
Creating directory: '/tmp/lite/target/install/1a96628d2b1833136b9540261d7cbe9e'
lite: Trampoline to 'lite' build system
lite: Updating manifest
lite: Install
/tmp/lite/target/install/1a96628d2b1833136b9540261d7cbe9e/usr/bin/lite: ELF 64-bit LSB
   executable, x86-64, version 1 (SYSV), dynamically linked (uses shared libs),
   for GNU/Linux 3.3.4, not stripped
\end{verbatim}
\caption{x86-64}\label{toolchain-usage:x86-64}
\hrulefill
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\hrulefill
\begin{verbatim}
pt mips-be_linux.3.3.4_gcc.4.6.3_eglibc_2.15-linux-gnu

lite: Creating directory: '/tmp/lite/target/lite/2db80f978ab6e8f96c1f7eca6c0fdf6b/build'
lite: Creating directory: '/tmp/lite/target/lite/2db80f978ab6e8f96c1f7eca6c0fdf6b/destdir'
Creating directory: '/tmp/lite/target/install/582598746ec21b53bdda1ba2874acc68'
lite: Trampoline to 'lite' build system
lite: Updating manifest
lite: Install
/tmp/lite/target/install/582598746ec21b53bdda1ba2874acc68/usr/bin/lite: ELF 32-bit MSB
   executable, MIPS, MIPS-I version 1 (SYSV), dynamically linked (uses shared libs),
   for GNU/Linux 3.3.4, with unknown capability 0x41000000 = 0xf676e75,
   with unknown capability 0x10000 = 0x70401, not stripped
\end{verbatim}
\caption{Big Endian Arm}\label{toolchain-usage:be-mips}
\hrulefill
\end{figure}

\begin{figure}
\hrulefill
\begin{verbatim}
pt mipsel-le_linux.3.3.4_gcc.4.6.3_eglibc_2.15-linux-gnu

lite: Creating directory: '/tmp/lite/target/lite/0831580102e7c193d2be1f39f8368f93/build'
lite: Creating directory: '/tmp/lite/target/lite/0831580102e7c193d2be1f39f8368f93/destdir'
Creating directory: '/tmp/lite/target/install/bff139f7e5b28e0213ecc0b9760829a9'
lite: Trampoline to 'lite' build system
lite: Updating manifest
lite: Install
/tmp/lite/target/install/bff139f7e5b28e0213ecc0b9760829a9/usr/bin/lite: ELF 32-bit LSB
   executable, MIPS, MIPS-I version 1 (SYSV), dynamically linked (uses shared libs),
   for GNU/Linux 3.3.4, with unknown capability 0xf41 = 0x756e6700,
   with unknown capability 0x70100 = 0x1040000, not stripped
\end{verbatim}
\caption{Little Endian MIPS}\label{toolchain-usage:le-mips}
\hrulefill
\end{figure}
\end{landscape}



\section{Toolchain Variables}\label{toolchain-usage:variables}

\lmsbw will set the \makefile variables to reference the configured
toolchain's versions of the following programs:

\begin{description}
\item{ADDR2LINE}
\item{AR}
\item{AS}
\item{CC}
\item{CPP}
\item{CXX}
\item{GCOV}
\item{GXX}
\item{LD}
\item{NM}
\item{OBJCOPY}
\item{OBJDUMP}
\item{POPULATE}
\item{RANLIB}
\item{READELF}
\item{SIZE}
\item{STRIP}
\end{description}


